---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  dpi = 300,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  warning = FALSE,
  message = FALSE 
)
```

# nzffdr

<!-- badges: start -->
<!-- badges: end -->

Warning, currently a WIP, will likely be code breaking changes occurring, until v 1.0.0!

The purpose of this package is to generate Stochastic Branching Networks (SBNs), which are frequently used to represent the branching structure of rivers in ecological models. The functions in the package rely heavily on the [igraph](https://igraph.org/r/) package. 

The fundamental unit of the SBN package is a downstream directed igraph object. The package contains functions for generating and manipulating these networks. THe functions *should* also work on river networks generated in other packages, which have subsequently been converted to [igraph](https://igraph.org/r/) objects (e.g. [OCNs](https://cran.r-project.org/web/packages/OCNet/vignettes/OCNet.html)).


### Installation
SBN is currently only accessible the via GitHub.
```{r, warning = FALSE, eval = FALSE}
devtools::install_github("flee598/SBN")
```


### Generatting SBNs
The SBNs generated are 


### Basic manipulation
The SBN package has a handful of functions for basic manipulation of networks.



### Calculating Strahler order
We can quickly calculate the Strahler order of vertices in a network. 


### A note on plotting SBNs
For quick plotting I have generally used [ggraph](https://www.data-imaginist.com/2017/ggraph-introduction-layouts/) with the `tree` layout. As far as I can tell, the `tree` algorithm doesn't play nice with a downstream weighted network, and therefore for plotting purposes you need to reverse the network to an upstream weighted one and it works fine. It is a bit fiddly, but I haven't found any simple "out-of-the-box" alternative as of yet. Another option is to generate your own vertex coordinates (as OCNet does), but I haven't gotten around to implementing that.
